---
title: "Analysis of  spotify playlist "
author: 
  - Daniel Xu
  - Dennis Netchitailo
  - Maria Mangru
thanks: "Code and data are available at: [TEMPNAME](https://github.com/possibleburger2/WHUNGEr)."
date: today
date-format: long
abstract: "This document analyzes something on spotify."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(spotifyr)
library(dplyr)
library(ggplot2)
library(tidyverse)

#Use IDs stored on system to access api
access_token <- get_spotify_access_token()

#Load artist information
bruno_mars <- get_artist_audio_features(artist = 'bruno mars')
lady_gaga <- get_artist_audio_features(artist = 'adele')

saveRDS(bruno_mars, "bruno_mars.rds")
saveRDS(lady_gaga, "lady_gaga.rds")

```

# Introduction
This study uses R packages [@citeR] to clean and analyze the dataset, including libraries from haven [@citehaven], dplyr [@citedplyr], readr [@citereadr], kableExtra [@citeknitr], and ggplot2 [@citeggplot]. The data we used is from IPUMS [@citedata].



# Estimates and the Actual Number of Respondents

```{r}
#| label: fig-data-summary
#| fig-cap: Bruno Mars lady gaga valence comparison graph
#| echo: false
rbind(lady_gaga, bruno_mars) |>
  select(artist_name, album_release_date, valence) |>
  mutate(album_release_date = ymd(album_release_date)) |> 
  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Valence",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@fig-data-summary shows comparison of blah blah

\newpage

\appendix

# Appendix {-}

# Instructions on How to Obtain the Data


\newpage

# References